---
title: "p8105_hw3_rac2224"
author: "Rachel Crowley"
date: "10/13/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r}
library(tidyverse)
library(p8105.datasets)
data("instacart")
```

Describing the Dataset "instacart"

Dataset size and structure 

* The number of observations in the dataset instacart is `r nrow(instacart)`. The number of columns in the dataset instacart is `r ncol(instacart)`. The variables in the dataset instacart are `r names(instacart)`. The dimensions of the dataset instacart are `r dim(instacart)`. 
* The structure of the variables within the dataset instacart are as follows: order_id is an `r typeof(pull(instacart, order_id))`, product_id is an `r typeof(pull(instacart, product_id))`, add_to_cart_order is an `r typeof(pull(instacart, add_to_cart_order))`, reordered is an `r typeof(pull(instacart, reordered))`, user_id is an `r typeof(pull(instacart, user_id))`, eval_set is a `r typeof(pull(instacart, eval_set))`, order_number is an `r typeof(pull(instacart, order_number))`, order_dow is an `r typeof(pull(instacart, order_dow))`, order_hour_of_day is an `r typeof(pull(instacart, order_hour_of_day))`, days_since_prior_order is an `r typeof(pull(instacart, days_since_prior_order))`, product_name is a `r typeof(pull(instacart, product_name))`, aisle_id is an `r typeof(pull(instacart, aisle_id))`, department_id is an `r typeof(pull(instacart, department_id))`, aisle is a `r typeof(pull(instacart, aisle))`, and department is a `r typeof(pull(instacart, department))`.

Key variable description

* The min, max, mean, and median of each of the integer variables are as follows: for order_ID, the minimum is `r min(pull(instacart, order_id))`, the maximum is `r max(pull(instacart, order_id))`, the mean is `r mean(pull(instacart, order_id))`, and the median is `r median(pull(instacart, order_id))`; for product_ID, the minimum is `r min(pull(instacart, product_id))`, the maximum is `r max(pull(instacart, product_id))`, the mean is `r mean(pull(instacart, product_id))`, and the median is `r median(pull(instacart, product_id))`; for add_to_cart_order, the minimum is `r min(pull(instacart, add_to_cart_order))`, the maximum is `r max(pull(instacart, add_to_cart_order))`, the mean is `r mean(pull(instacart, add_to_cart_order))`, and the median is `r median(pull(instacart, add_to_cart_order))`; for reordered, the minimum is `r min(pull(instacart, reordered))`, the maximum is `r max(pull(instacart, reordered))`, the mean is `r mean(pull(instacart, reordered))`, and the median is `r median(pull(instacart, reordered))`; for user_id, the minimum is `r min(pull(instacart, user_id))`, the maximum is `r max(pull(instacart, user_id))`, the mean is `r mean(pull(instacart, user_id))`, and the median is `r median(pull(instacart, user_id))`; for order_number, the minimum is `r min(pull(instacart, order_number))`, the maximum is `r max(pull(instacart, order_number))`, the mean is `r mean(pull(instacart, order_number))`, and the median is `r median(pull(instacart, order_number))`; for order_dow, the minimum is `r min(pull(instacart, order_dow))`, the maximum is `r max(pull(instacart, order_dow))`, the mean is `r mean(pull(instacart, order_dow))`, and the median is `r median(pull(instacart, order_dow))`; for order_hour_of_day, the minimum is `r min(pull(instacart, order_hour_of_day))`, the maximum is `r max(pull(instacart, order_hour_of_day))`, the mean is `r mean(pull(instacart, order_hour_of_day))`, and the median is `r median(pull(instacart, order_hour_of_day))`; for days_since_prior_order, the minimum is `r min(pull(instacart, days_since_prior_order))`, the maximum is `r max(pull(instacart, days_since_prior_order))`, the mean is `r mean(pull(instacart, days_since_prior_order))`, and the median is `r median(pull(instacart, days_since_prior_order))`; for aisle_id, the minimum is `r min(pull(instacart, aisle_id))`, the maximum is `r max(pull(instacart, aisle_id))`, the mean is `r mean(pull(instacart, aisle_id))`, and the median is `r median(pull(instacart, aisle_id))`; and for department_id, the minimum is `r min(pull(instacart, department_id))`, the maximum is `r max(pull(instacart, department_id))`, the mean is `r mean(pull(instacart, department_id))`, and the median is `r median(pull(instacart, department_id))`. 
* The number of missing observations for each variable are as follows: for order_id is `r length(which(is.na(instacart$order_id)))`, for product_id is `r length(which(is.na(instacart$product_id)))`, for add_to_cart_order is `r length(which(is.na(instacart$add_to_cart_order)))`, for reordered is `r length(which(is.na(instacart$reordered)))`, for user_id is `r length(which(is.na(instacart$user_id)))`, for eval_set is `r length(which(is.na(instacart$eval_set)))`, for order_number is `r length(which(is.na(instacart$order_number)))`, for order_dow is `r length(which(is.na(instacart$order_dow)))`, for order_hour_of_day is `r length(which(is.na(instacart$order_hour_of_day)))`, for days_since_prior_order is `r length(which(is.na(instacart$days_since_prior_order)))`, for product_name is `r length(which(is.na(instacart$product_name)))`, for aisle_id is `r length(which(is.na(instacart$aisle_id)))`, for department_id is `r length(which(is.na(instacart$department_id)))`, for aisle is `r length(which(is.na(instacart$aisle)))`, and for department is `r length(which(is.na(instacart$department)))`.

Examples of observations

* The mean order hour of day for products from the produce department is `r mean(subset(instacart, department=='produce')$order_hour_of_day)`, the mean order hour of day for products from the frozen department is `r mean(subset(instacart, department=='frozen')$order_hour_of_day)`, and the mean order hour of day for products from the household department is `r mean(subset(instacart, department=='household')$order_hour_of_day)`. The largest number of days since prior order for an order that included something from the eggs aisle is `r max(subset(instacart, aisle =='eggs')$days_since_prior_order)` and for an order that included something from the coffee aisle is `r max(subset(instacart, aisle =='coffee')$days_since_prior_order)`.



